---
title: "Project Template for STAT 5385"
lang: en
output: word_document
---

```{r "Load libraries", echo=FALSE, message=FALSE, warning=FALSE}

```

```{r "Load data"}

```

```{r "Munge data"}

```

resources for the background
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8692088/

# Abstract
(a) Describe study purpose
(b) Provide balanced summary of what was done and what was found

# Introduction

## Background/rationale

Explain the scientific background and rationale for the investigation being reported

## Objectives	
State specific objectives, including any prespecified hypotheses

## Setting
Describe the setting, locations, and relevant dates, including periods of recruitment, exposure, follow-up, and data collection

## Participants	
Describe participants here

## Variables
Clearly define all outcomes, exposures, predictors, potential confounders, and effect modifiers. Give diagnostic criteria, if applicable

## Data sources/ measurement
For each variable of interest, give sources of data and details of methods of assessment (measurement). Describe comparability of assessment methods if there is more than one group

## Bias
Describe any efforts to address potential sources of bias

## Study size
Explain how the study size was arrived at


## Statistical methods
Describe all statistical methods, including those used to control for confounding  

# Results

## Descriptive data	
Give characteristics of study participants available, information on exposures and potential confounders, and any missing data


## Main results	
Describe in detail results and provide in context description

## Other analyses	
Report other analyses done—eg analyses of subgroups and interactions, and sensitivity analyses

# Discussion

## Key results
Summarise key results with reference to study objectives

## Limitations
Discuss limitations of the study, taking into account sources of potential bias or imprecision. Discuss both direction and magnitude of any potential bias

## Interpretation
Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence

## Generalisability
Discuss the generalisability (external validity) of the study results
