---
title: "Lab - 1.45"
author: "STAT 5385"
date: "1/20/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(summarytools)
library(dplyr)

```

## SENIC Data Read-in

```{r}

setwd("~/teaching/regression analysis/labs")

senic=read.table("C:/Users/awagler2/Documents/teaching/regression analysis/Data Sets/Appendix C Data Sets/APPENC01.txt")
colnames(senic)=c("ID","LOS","Age","Infec","Cul","Xray","beds","Med","region","avg","nurses","fands")
head(senic)

senic %>% dfSummary() %>% view()
```

## Estimate the models (part a of 1.45)

For infection rate the estimated model is $\hat{y} = 6.34 + 0.76 x$ where x is infection rate.  Our MSE = 2.638. There is a positive linear relationship between infection rate and LOS.

```{r}
#lm stands for linear model=regression model
mod.a=lm(LOS~Infec,data=senic)
summary(mod.a)
#provides MSE, residual variance
anova(mod.a)
#don't do this until chap 3
#plot(mod.a)

mod.b=lm(LOS~fands,data=senic)
summary(mod.b)
anova(mod.b)

mod.c=lm(LOS~Xray,data=senic)
summary(mod.c)
anova(mod.c)
```

## Plot the model

For this section comment on whether you think the model is linear or not.

```{r}
with(senic,plot(Infec,LOS))
abline(lm(LOS~Infec,data=senic))

with(senic,plot(fands,LOS))
abline(lm(LOS~fands,data=senic))

with(senic,plot(Xray,LOS))
abline(lm(LOS~Xray,data=senic))

```

## This is 1.46 more or less-see if you can finish it for regions 2,3 and 4 (this is completely optional)

```{r}
mod.r1=lm(LOS~Infec,data=senic[senic$region==1,])
summary(mod.r1)

mod.r2=lm(LOS~Infec,data=senic[senic$region==2,])
summary(mod.r2)

mod.r3=lm(LOS~Infec,data=senic[senic$region==3,])
summary(mod.r3)

mod.r4=lm(LOS~Infec,data=senic[senic$region==4,])
summary(mod.r4)

```
