---
title: 'STAT 5385: Lab `r params$lab_number`'
author: "Willliam Ofosu Agyapong"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    number_section: yes
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \rhead{William O. Agyapong}
  - \lhead{STAT 5385 - Lab `r params$lab_number`}
  - \cfoot{\thepage}
geometry: margin = 0.8in
fontsize: 10pt
params:
  lab_number: 4
---


```{r setup, include=FALSE}
# Set global options for output rendering
knitr::opts_chunk$set(echo = T, warning = F, message = F, fig.align = "center")

# Load required packages
# library(summarytools)
library(dplyr)
library(tidyr)
library(purrr)
library(broom)
library(ggplot2)
# library(kableExtra)
library(knitr)
library(patchwork)

# Set the current working directory to the file path
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) 

# Set default rounding to 4 decimal places
options(digits = 4)

# Define a default ggplot theme
theme_set(theme_classic())
```

# Introducing the SENIC data
```{r, echo=FALSE}
var_table <- data.frame(
  actual = c("Identification Number", "Length of stay", "Age", "Infection risk",
             "Routine culturing ratio", "Routine chest X-ray ratio",
             "Number of beds", "Medical school affiliation", "Region",
             "Average daily census", "Number of nurses", "Available facilities and services"),
  coded_names = c("ID", "LOS", "age", "infec_risk", "cul_ratio", "xray_ratio",
               "beds", "med", "region", "ADC", "nurses", "AFS")
)

kable(var_table, col.names = c("Variable Name", "Coded As"),
      caption = "Variables in the SENIC dataset")
```

## Reading in the dataset

```{r}
# Import the dataset from local drive
senic <- read.table(file = "../Data Sets/Appendix C Data Sets/APPENC01.txt",
                    header = FALSE)

# Rename columns
names(senic) <- var_table$coded_names

# View first few observations; Everything looks good.
head(senic)

# Obtain summary report
# senic %>% dfSummary() %>% view()
```


# Problem 3.27

## Part (a): Diagnostic Plots

### Infection Risk

```{r}
# Fit a linear model
mod_infec_risk <- lm(LOS ~ infec_risk, data = senic)

# residuals against infection risk
 par(mfrow=c(2,2))

plot(senic$infec_risk, resid(mod_infec_risk),
     xlab = "Infection Risk (X)", ylab = "Residual"); abline(h=0)

plot(senic$infec_risk, rstudent(mod_infec_risk),
     xlab = "Infection Risk (X)", ylab = "Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_infec_risk, which = 2)

# investigating outliers
summary(senic[, c("LOS", "infec_risk")])

# look at potential outliers: 47 and 112
senic[c(47,76,112), c("LOS", "infec_risk")]
```

- Potential outliers, Looking at the residual versus the majority of observations are between 4 and 6 rate of infection risk.
They can influence the kind of transformation

- 

### Available Facilities and Services (AFS)

```{r}
mod_AFS <- lm(LOS ~ AFS, data = senic)
# residuals agains infection risk
par(mfrow = c(2,2))
plot(senic$AFS, resid(mod_AFS),
     xlab = "AFS (X)", ylab = "Residual"); abline(h=0)

plot(senic$AFS, rstudent(mod_AFS),
     xlab = "AFS (X)", ylab = "Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_AFS, which = 2)

# investigating outliers
summary(senic[, c("LOS", "AFS")])

# look at potential outliers: 47 and 112
senic[c(34,47,112), c("LOS", "AFS")]
```


### Chest X-ray Ratio


```{r}
mod_xray_ratio <- lm(LOS ~ xray_ratio, data = senic)
# residuals agains infection risk
par(mfrow = c(2,2))
plot(senic$xray_ratio, resid(mod_xray_ratio),
     xlab = "X-ray Ratio", ylab = "Residual"); abline(h=0)

plot(senic$xray_ratio, rstudent(mod_xray_ratio),
     xlab = "X-ray Ratio (X)", ylab = "Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_xray_ratio, which = 2)

# investigating outliers
summary(senic[, c("LOS", "xray_ratio")])

# look at potential outliers: 47 and 112
senic[c(33,47,112), c("LOS", "xray_ratio")]
```

- Two data points at the 47th and 112th observations were identified as obvious outliers in all three models.
- The two extreme observations appear to be the main factor contributing to the of lack of fit in all the models.
- Therefore, except for the outliers causing some departures, a linear regression appears to be slightly more appropriate in the cases involving X-ray ratio and AFS as predictors than the model involving infection risk.


## Part (b): Models without outlying observations

Observations **47** and **112** appeared to be extreme values for all the three models involving the three individual predictors. As a next step, we refit all models without these two observations  to assess their impact on model performance.

### Infection Risk
```{r}
senic2 <- senic[-c(47, 112),] # data without outliers
mod_infec_risk_b <- lm(LOS ~ infec_risk, data = senic2)
# summary(mod_infec_risk_b)

# residuals agains infection risk
par(mfrow = c(2,2))
plot(senic2$infec_risk, resid(mod_infec_risk_b),
     xlab ="Infection Risk (X)",ylab ="Residual"); abline(h=0)

plot(senic2$infec_risk, rstudent(mod_infec_risk_b),
     xlab ="Infection Risk (X)",ylab ="Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_infec_risk_b, which = 2)
# investigating outliers
# look at 
summary(senic[, c("LOS","infec_risk")])
senic2[c(34,104), c("LOS","infec_risk")]

```

#### $95\%$ Prediction Intervals
```{r}
# obtaining 95% prediction interval at Xh = 6.5
kable(predict(mod_infec_risk_b, newdata = data.frame(infec_risk = 6.5), interval = "prediction"))

# obtaining 95% prediction interval at Xh = 5.9
kable(predict(mod_infec_risk_b, newdata = data.frame(infec_risk = 5.9), interval = "prediction"))
```
**Clearly, observations $Y_{47} = 19.56$ and $Y_{112} = 17.94$ lie far outside the two prediction intervals corresponding to $X_h=6.5$ and $X_h=5.9$, respectively.**

The fact that the observations fall outside the prediction intervals could be an indication that those observations most likely arose from measurement errors.

### Available Facilities and Services (AFS)
```{r}
senic2 <- senic[-c(47, 112),] # data without outliers
mod_AFS_b <- lm(LOS ~ AFS, data = senic2)
# summary(mod_AFS_b)

# residuals agains infection risk
par(mfrow = c(2,2))
plot(senic2$AFS, resid(mod_AFS_b),
     xlab ="AFS (X)",ylab ="Residual"); abline(h=0)

plot(senic2$AFS, rstudent(mod_AFS_b),
     xlab ="AFS (X)",ylab ="Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_AFS_b, which = 2)
# investigating outliers
# look at 
summary(senic[, c("LOS","AFS")])
senic2[c(34,104), c("LOS","AFS")]

```


#### $95\%$ Prediction Intervals
```{r}
# obtaining 95% prediction interval at Xh = 51.4
kable(predict(mod_AFS_b, newdata = data.frame(AFS = 51.4), interval = "prediction"))

# obtaining 95% prediction interval at Xh = 62.9
kable(predict(mod_AFS_b, newdata = data.frame(AFS = 62.9), interval = "prediction"))
```
**The same observation is made here. The observations fall outside the prediction  intervals**.

### Chest X-ray Ratio
```{r}
senic2 <- senic[-c(47, 112),] # data without outliers
mod_xray_ratio_b <- lm(LOS ~ xray_ratio, data = senic2)
# summary(mod_AFS_b)

# residuals agains infection risk
par(mfrow = c(2,2))
plot(senic2$xray_ratio, resid(mod_xray_ratio_b),
     xlab ="X-ray Ratio",ylab ="Residual"); abline(h=0)

plot(senic2$xray_ratio, rstudent(mod_xray_ratio_b),
     xlab ="X-ray Ratio (X)",ylab ="Studentized Residual"); abline(h=0)

# normal probability plot
plot(mod_xray_ratio_b, which = 2)
# investigating outliers
# look at 
summary(senic2[, c("LOS","xray_ratio")])
senic2[c(34,104), c("LOS","xray_ratio")]

```

#### $95\%$ Prediction Intervals
```{r}
# obtaining 95% prediction interval at Xh = 113.7
kable(predict(mod_xray_ratio_b, newdata = data.frame(xray_ratio = 113.7), interval = "prediction"))

# obtaining 95% prediction interval at Xh = 91.8
kable(predict(mod_xray_ratio_b, newdata = data.frame(xray_ratio = 91.8), interval = "prediction"))
```
**As noted earlier in the other cases, the observations are also outside these prediction intervals, and  this could be a signal that the observations are that extreme due to possible measurement errors**

**When comparing the models**


# Problem 3.28

```{r collapse=FALSE}
models <- list() # initialize empty list

# Get the various regions
regions <- as.integer(levels(as.factor(senic$region)))

# fit individual models for each region
for(i in regions) {
  models[[i]] = senic %>%
filter(region == i) %>%
    lm(LOS ~ infec_risk, data = .)
  # print(paste("Estimated model summary for region", i))
  # print(summary(models[[i]]))
}
 
```


## Diagnostic plots

We look at diagnostic plots for models involving Length of Stay (Y) and Infection Risk for each region.

### Region 1
```{r}

# region 1
par(mfrow = c(1,2))
plot(senic[senic$region==1,]$infec_risk, resid(models[[1]]),
     xlab ="Infection Risk (for region 1)",ylab = "Residual"); abline(h=0) # residuals against infection risk
plot(models[[1]], which = 2)# normal probability plot
```


### Region 2
```{r}
par(mfrow = c(1,2))
plot(senic[senic$region==2,]$infec_risk, resid(models[[2]]),
     xlab ="Infection Risk (for region 2)",ylab = "Residual"); abline(h=0) # region 2
plot(models[[2]], which = 2)# normal probability plot
```


### Region 3
```{r}
par(mfrow = c(1,2))
plot(senic[senic$region==3,]$infec_risk, resid(models[[3]]), 
     xlab ="Infection Risk (for region 3)",ylab = "Residual"); abline(h=0) # region 3
plot(models[[3]], which = 2)# normal probability plot
```



### Region 4
```{r}
par(mfrow = c(1,2))
plot(senic[senic$region==4,]$infec_risk, resid(models[[4]]),
     xlab ="Infection Risk (for region 4)",ylab = "Residual"); abline(h=0) # region 4
plot(models[[4]], which = 2)# normal probability plot

```


The residual versus predictor plots suggest that the error variance differ substantially across the four regions. There appear to be increasing error variance with levels of infection risk in region 1, while the other regions exhibit randomness.

Additionally, except region 2 there appear to be serious departure of error term distribution from normality.
